---
title: "Assignment 1"
author: "Levi Uy"
format: html
---

**Overview:** *For this assignment, you are going to download some variables using the `wbstats` and `vdemdata` packages. Then you are going to wrangle these data and merge the two data sets into one and analyze how they relate to women's representation in national parliaments. Do your work for each step in the code chunks provided. Be sure to label your code chunks.*

## Step 1: Download data from the V-Dem Dataset

*Look at the V-Dem [codebook](https://www.v-dem.net/data/reference-documents/). Identify two measures of democracy that are not the polyarchy score that we used in Module 1.2.*

*Try to pick variables that will relate to women's representation and read the description of the variable to make sure that it includes data for a sufficient number of years, e.g. that the data will be somewhat recent. Select the most recent 20 years of data for your analysis.*

*Make sure to load the packages that you need and `glimpse()` the data or `View()` it to make sure that it downloaded properly.*

```{r}
#| label: v_dem

library(vdemdata)
library(dplyr)

vdem_ind <- vdem |>
  filter(year >= 2002) |>
  select(
    country = country_name,
    vdem_ctry_id = country_id,
    year,
    egal = v2x_egaldem,
    civil = v2x_cspart,
    region = e_regionpol_6C
  ) |>
  mutate(
    region = case_match(region, 
                        1 ~ "Eastern Europe",
                        2 ~ "Latin America",
                        3 ~ "Middle East",
                        4 ~ "Africa",
                        5 ~ "The West",
                        6 ~ "Asia")
  )
```

## Step 2: Download data from the World Bank

*Next, download the variable on women's represenation that we used in Module 1.2 ("SG.GEN.PARL.ZS") and at least one additional measure related to women's empowerment. Go to the [WDI](https://databank.worldbank.org/source/world-development-indicators) site the wb_search() function to identify relevant variables. Download the most recent 20 years of data for your analysis.*

*Make sure that the indicator has enough data to conduct your analysis, i.e. that it has data for most countries and years.*

```{r}
#| label: wb_stats

library(wbstats)
library(janitor)

#women_bus <- wb_search("women business")

#print(women)bus, n=26)

# Storing variables in indicator object

indicators <- c("women_rep" = "SG.GEN.PARL.ZS", "women_bus" = "SG.LAW.INDX")

# Downloading the data

women_emp <- wb_data(indicators, mrv = 20) |>
  select(!iso2c) |>
  rename(year = date) |>
  mutate(
    women_bus = round_to_fraction(women_bus, denominator = 100),
    women_rep = round_to_fraction(women_rep, denominator = 100)
  )

glimpse(women_emp)

```

## Step 3: Merge the data

*Now add country codes using the `countrycode` package and merge the data using `left_join()`.*

```{r}
#| label: country_codes

# Adding country_codes

library(countrycode)

vdem_var <- vdem_ind |>
  mutate(iso3c = countrycode(
    sourcevar = vdem_ctry_id,
    origin = "vdem",
    destination = "wb", 
  )) |>
  relocate(iso3c, .after = vdem_ctry_id)

# View the data
# glimpse(vdem_vars)
```

```{r}
#| label: join

library(readr)

women_dem <- left_join(vdem_var, women_emp, by = c("iso3c", "year")) 


# View the data
glimpse(women_dem)
```

## Step 4: Summarize your combined data set

*Use `group_by()`, `summarize()` and `arrange()` to glean insights about your data. For example, how do regions compare on mean values of women's representation and how do these values relate to the values of other variables in your data set? Which countries stand out in terms of women's representation? Etc.*

```{r}
#| label: Group by(), summarize(), and arrange()

dem_summary <- women_dem |> 
  group_by(country.x)  |> 
  summarize(           
    egal = mean(egal, na.rm = TRUE), 
    civil = mean(civil, na.rm = TRUE), 
    women_rep = mean(women_rep, na.rm = TRUE), 
    women_bus = mean(women_bus, na.rm = TRUE),
  ) |> 
  arrange(country.x) 

# Data insights
# The West has the highest mean values for women's representation in parliament with a value of 29.83. Latin America ranks second with 23.55, and then Eastern Europe with a value of 19.75. The Middle East ranks last with a value of 11.57 in women's representation. These rankings are mirrored in the egalitarian democracy index from the V-Dem dataset. The egalitarian index considers 1) the equal protection of rights and freedoms - and 2) the equal distribution of resources - across all social groups. The West ranks first in this index with a value of 0.78 (the index measures on a scale of 0 to 1), Latin America and Eastern Europe both see a mean value of .43. The Middle East ranks last with a value of .21. Again, a similar ranking is seen in the civil society participation index from the V-Dem dataset. This variable considers how robust civil society is, and how much consideration these groups are given in terms of policy decisions (the index measures on a scale of 0 to 1). The West ranks first with .92, Latin America second with .72, Africa third with .67, and the Middle East last with .44. Despite The West ranking first in women's representation, Rwanda (Africa) ranks first with 57.38, Sweden (The West) second with 45.60, and Cuba (Latin America) third with 45.53. In contrast, Yemen ranks last with a value of 0.23. However, it would be interesting to analyze other countries that did not have sufficient data. These countries include Hong Kong, Palestine, Taiwan, etc. 
```

## Step 5: Interpret your data

*Write a paragraph or so relating your insights to the Norris reading about electoral reform and women's representation. Picking a handful of country cases from your data, how does your analysis relate to her arguments about the relevance of electoral systems, statutory quotas, reserved seats or voluntary quotas?*

Although Norris discusses the effect of proportional electoral systems leading to more women's representation, the electoral system is not a strong dependent factor of other empowerment indicators. For example, Costa Rica has a proportional electoral system and reports .81 on the egalitarian index, .94 on the civil society index, 38.68 on the women's representation index, and 80.75 for the business and law index (which measures how laws/regulations impact economic opportunity for women out of a scale of 1 to 100). However, Mexico has a combined electoral system and reports .62 on the civil society index, 35.07 on the women's representation index, and 81 on the business and law index. Moreover, France has a majoritarian system and reports .78 on the egalitarian index, .91 on the civil society index, 25.43 on the women's representation index, and 97 on the business and law index. Although Costa Rica has the highest value for the representation index, the electoral system does not seem to be a big determinant of other empowerment indicators for women.

Norris also acknowledges that various factors contribute to the effectiveness of statutory quotas, which can be seen among general factors of women's representation and empowerment. For example, Norris reports that Afghanistan reserves 27% of seats for women. However, Afghanistan reports a mean of .19 on the egalitarian index, 31.21 for the business and law index, and 27.56 for women's representation. Although Afghanistan saves the highest percentage of seats out of the countries Norris provides, they report relatively low scores in other empowerment indicators. Pakistan is another example of saving a relatively high percentage of seats for women, but reports low scores in other indicators. For example, they report a mean value of .15 for the egalitarian index, 21.16 for women's representation, and 45.62 for women's business and law index. These findings support the argument that there are many other considerations when evaluating the impact of statutory quotas.
